Depression and suicidality as credible signals of need

### 2021-05-07 Paper.rmd
Added discussion of problems with pilot study control condition

<!-- Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are consistent with this hypothesis. In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception and the number of defectors within the broader population [@Rosenthal1993], with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact that depression and suicide attempts become relatively less costly to one's fitness prospects with increased need [@Hagen2003], suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose. -->

@@

Another limitation is that our control included a signal in the form of a verbal request, preventing comparisons to non-signaling strategies. Similarly, our design did not include vignettes with male signalers. For this reason, we have no data on the possibility of sex differences in the effectiveness of the signaling strategies examined. Although not predicted theoretically, such differences are possible if the costs of signaling vary between the sexes due to differential access to alternative bargaining strategies [@HagenRosenstrom2016] or if one sex tends to suffer greater negative reputational effects  when displaying the emotions and behaviors in the vignettes. It is also possible the costliness of the situations presented in the vignettes differ by sex.

@@

<!-- Slider anchor  -->
<!-- Regex: stopped in SI -->
<!-- T\d \w+ -->
# Conclusion

Our results suggest that the costs of depression and suicidality, which in mainstream theories are taken as evidence of brain dysfunction [@SymeHagen2021], might instead function to help victims elicit support from reluctant social partners with whom they are in conflict. This coupled with findings that depression can simultaneously elicit both positive and negative responses [@HokansonEtAl1986; @stephens1987], suggests that more attention should be paid to the potential for positive responses towards depressed individuals, especially from individuals who normally receive substantial benefits from the depressed or suicidal individual.

 <!-- The need to explain suicide's connection to depression is often overlooked by explanations for major depression, particularly those which propose non-signaling functions. One explanation for their connection, is that both suicidality and depression may function to limit negative fitness outcomes in times of substantial adversity [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], with costlier signals being required to elicit support when the payoffs to faking need are high enough so that less costly signals are not reliable indicators of one's need [@Rosenthal1993]. For this hypothesis to be correct, instrumental support would need to be provided frequently enough with large enough impacts so that the costs of signaling are outweighed by the benefits on average. In support of this, we found exposure to depression and suicide attempts resulted in greater belief in need and increased likelihood of support in fictional scenarios. -->

@@

Briefly, Participants read a vignette that put them in situations in which an imagined sister was requesting a loan of \$50,000 from the participant that was originally saved for the participant's imagined daughter's college education. The sister claims the money is needed for medical treatment for her own child, but she has incentives to deceive. The manipulations were signals very similar to those used in this study, and the outcomes were belief that the sister really needed the money and intention to loan her the money. See XX for the full write-up. See <https://osf.io/3rg8b> for preregistration of the pilot study. See <http://doi.org/10.5281/zenodo.4637883> for pilot data.

@@

We learned five major lessons from the MTurk pilot study that influenced the design of the current study. First, as predicted, belief in the sister's need was highest in the *Depression* condition and lowest in the *Verbal request* condition, giving us confidence to continue this research (see Figure \@ref(fig:piloteffects)). Second, to rule out that believability and willingness to help were simply artifacts of the sister's psychiatric distress, we included a "signaling" condition in which the sister exhibited schizophrenic symptoms. As predicted, believability and willingness to help in this condition were dramatically lower than in any other condition, ruling out this alternative explanation. We therefore did not include the schizophrenic condition in the current study. Third, baseline (T1) believability was very high (mean = `r signif(pilot_results$meanT1belief, 2)` and median = `r signif(pilot_results$medianT1belief, 2)` on a 0-100 point scale), leaving little scope for signals to increase believability. In the current study we therefore endeavored to provide details in the vignettes that would undermine the victim's believability at baseline. Fourth, contrary to our hypothesis, the *Suicide threat* and *Suicide attempt* signals reduced believability relative to *Depression*. We surmised that the sister's adversity -- a sick child -- was not sufficient to justify a suicide attempt, thus undermining her believability. In the current study the signaler herself is therefore the victim, and the forms of adversity -- physical and sexual assault and thwarted and forced marriage -- are those that are frequently associated with suicidality. Fifth, the content of the vignette matters, so instead of a single vignette we decided to use four different vignettes to help determine the extent to which signaling effects were sensitive to seemingly irrelevant details of the vignettes.

@@

```{r distaction, fig.width=10, fig.height=10, fig.cap="Histograms of *T2 Action* by signal and vignette."}

@@

```{r demoeffectsaction, fig.width=9, fig.height=8, fig.cap="Demographic effects on Action. **Top**: Association of age with *T2 Action*, controlling for *T1 Action*, sex, and vignette. **Bottom**: Association of sex with *T2 Action* controlling for *T1 Action*, age, and vignette. Estimated using a generalized linear regression model with the quasibinomial family."}
# effectsPlots$m25plotAge / effectsPlots$m25plotSex
```

```{r demovignettebelief, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with *T2 Belief* by vignette, controlling for *T1 Belief* and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
# (effectsPlots_US$m18plot / effectsPlots_US$m13plot)
```

```{r demovignetteaction, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with *T2 Action* by vignette, controlling for *T1 Action* and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
# effectsPlots_US$m19plot / effectsPlots_US$m15plot
```

@@

```{r mediationsuicide, fig.height=4, fig.cap="The effect of a Suicide attempt (treatment) vs. Depression (control) on Action, [0-1 scale], is only partially mediated by Belief. ACME: Average causal mediation effect. ADE: Average direct effect. The mediation model controlled for *T1 Belief*, and the outcome model controlled for *T1 Belief* and *T1 Action*. Both models were GLMs with the binomial family (the mediation package does not support the quasibinomial family)."}
ggmediation(med2)
```

@@

```{r divisionage, fig.width=9, fig.cap="Interaction of age with signals on *T2 Division* of dowry between the daughters in the thwarted marriage vignette."}
effectsPlots_India$m8plot
```

@@

```{r t3action, fig.width=10, fig.cap="The effect of post-signal (T3) information that the victim was telling the truth on the likelihood of acting to help her (*T3 Action*), controlling for *T2 Action*, with a term for the interaction of *T2 Action* and vignette."}
effectsPlots$m26plot + theme(panel.spacing = unit(1.5, "lines"))
```

@@

```{r preregbelief, fig.width=8, fig.height=8, fig.cap="The effect of the signals on *T2 Belief* (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1, xvar='signal',ylbl='T2 Belief', data=d)

@@

```{r preregbeliefB, fig.width=8, fig.height=8, fig.cap="The effect of the signals on *T2 Belief* (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively, with an interaction between T1 values and the signal. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1b, xvar='signal', by='T1Belief', ylbl='T2 Belief', data=d)

@@

```{r modelcompareplots, fig.width=11, fig.cap="Comparison of the coefficients and 95% CIs of the quasibinomial glm, fractional, and bootstrapped logistic regression models of *T2 Belief* and *T2 Action* (for details, see the Statistical methods section)."}

>>>>

Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are consistent with this hypothesis. In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception and the number of defectors within the broader population [@Rosenthal1993], with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact that depression and suicide attempts become relatively less costly to one's fitness prospects with increased need [@Hagen2003], suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose.

@@

Another limitation is that our control included a signal in the form of a verbal request, preventing comparisons to non-signaling strategies. Similarly, our design did not include vignettes with male signalers. For this reason, we have no data on the possibility of sex differences in the effectiveness of the signaling strategies examined. Although not predicted theoretically, such differences are possible if the costs of signaling vary between the sexes due to differential access to alternative bargaining strategies [@HagenRosenstrom2016] or if one sex suffered greater negative reputational effects due to emotions and behaviors in the vignettes. It is also possible the costliness of the situations presented in the vignettes differ by sex.

@@

Slider anchor 
Regex: stopped in SI
T\d \w+
# Conclusion

The need to explain suicide's connection to depression is often overlooked by explanations for major depression, particularly those which propose non-signaling functions. One explanation for their connection, is that both suicidality and depression may function to limit negative fitness outcomes in times of substantial adversity [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], with costlier signals being required to elicit support when the payoffs to faking need are high enough so that less costly signals are not reliable indicators of one's need [@Rosenthal1993]. For this hypothesis to be correct, instrumental support would need to be provided frequently enough with large enough impacts so that the costs of signaling are outweighed by the benefits on average. In support of this, we found exposure to depression and suicide attempts resulted in greater belief in need and increased likelihood of support in fictional scenarios. This coupled with findings that depression can simultaneously elicit both positive and negative responses [@HokansonEtAl1986; @stephens1987], suggests that more attention should be paid to the potential for positive responses towards depressed individuals, especially from individuals who normally receive substantial benefits from the depressed or suicidal individual.

@@

Briefly, participants read a vignette that put them in situations in which an imagined sister was requesting a loan of \$50,000 from the participant that was originally saved for the participant's imagined daughter's college education. The sister claims the money is needed for medical treatment for her own child, but she has incentives to deceive. The manipulations were signals very similar to those used in this study, and the outcomes were belief that the sister really needed the money and intention to loan her the money. See XX for the full write-up. See <https://osf.io/3rg8b> for preregistration of the pilot study. See <http://doi.org/10.5281/zenodo.4637883> for pilot data.

@@

We learned six major lessons from the MTurk pilot study that influenced the design of the current study. First, as predicted, belief in the sister's need was highest in the *Depression* condition and lowest in the *Verbal request* condition, giving us confidence to continue this research (see Figure \@ref(fig:piloteffects)). Second, to rule out that believability and willingness to help were simply artifacts of the sister's psychiatric distress, we included a "signaling" condition in which the sister exhibited schizophrenic symptoms. As predicted, believability and willingness to help in this condition were dramatically lower than in any other condition, ruling out this alternative explanation. We therefore did not include the schizophrenic condition in the current study. Third, baseline (T1) believability was very high (mean = `r signif(pilot_results$meanT1belief, 2)` and median = `r signif(pilot_results$medianT1belief, 2)` on a 0-100 point scale), leaving little scope for signals to increase believability. In the current study we therefore endeavored to provide details in the vignettes that would undermine the victim's believability at baseline. Fourth, contrary to our hypothesis, the *Suicide threat* and *Suicide attempt* signals reduced believability relative to *Depression*. We surmised that the sister's adversity -- a sick child -- was not sufficient to justify a suicide attempt, thus undermining her believability. In the current study the signaler herself is therefore the victim, and the forms of adversity -- physical and sexual assault and thwarted and forced marriage -- are those that are frequently associated with suicidality. Fifth, the content of the vignette matters, so instead of a single vignette we decided to use four different vignettes to help determine the extent to which signaling effects were sensitive to seemingly irrelevant details of the vignettes. Sixth, in a randomized controlled trial, the control condition should be identical to the treatment condition in every respect except for the factor being tested, which in this study was signal cost. Post hoc, we realized that the control condition in the pilot study differed substantially from the treatment condition beyond signal cost: in the treatment conditions, participants encountered the sister twice: once when she requests help, and a second time when she signals. In our faulty control, participants only encountered the sister once, when she requests help, with no second encounter. In this study we therefore ensured that the control condition (Verbal request) was identical to the treatment conditions except for signal cost.

@@

```{r distaction, fig.width=10, fig.height=10, fig.cap="Histograms of T2 Action by signal and vignette."}

@@

```{r demoeffectsaction, fig.width=9, fig.height=8, fig.cap="Demographic effects on Action. **Top**: Association of age with T2 Action, controlling for T1 Action, sex, and vignette. **Bottom**: Association of sex with T2 Action controlling for T1 Action, age, and vignette. Estimated using a generalized linear regression model with the quasibinomial family."}
# effectsPlots$m25plotAge / effectsPlots$m25plotSex
```

```{r demovignettebelief, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Belief by vignette, controlling for T1 Belief and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
# (effectsPlots_US$m18plot / effectsPlots_US$m13plot)
```

```{r demovignetteaction, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Action by vignette, controlling for T1 Action and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
# effectsPlots_US$m19plot / effectsPlots_US$m15plot
```

@@

```{r mediationsuicide, fig.height=4, fig.cap="The effect of a Suicide attempt (treatment) vs. Depression (control) on Action, [0-1 scale], is only partially mediated by Belief. ACME: Average causal mediation effect. ADE: Average direct effect. The mediation model controlled for T1 Belief, and the outcome model controlled for T1 Belief and T1 Action. Both models were GLMs with the binomial family (the mediation package does not support the quasibinomial family)."}
ggmediation(med2)
```

@@

```{r divisionage, fig.width=9, fig.cap="Interaction of age with signals on T2 Division of dowry between the daughters in the thwarted marriage vignette."}
effectsPlots_India$m8plot
```

@@

```{r t3action, fig.width=10, fig.cap="The effect of post-signal (T3) information that the victim was telling the truth on the likelihood of acting to help her (T3 Action), controlling for T2 Action, with a term for the interaction of T2 Action and vignette."}
effectsPlots$m26plot + theme(panel.spacing = unit(1.5, "lines"))
```

@@

```{r preregbelief, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1, xvar='signal',ylbl='T2 Belief', data=d)

@@

```{r preregbeliefB, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively, with an interaction between T1 values and the signal. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1b, xvar='signal', by='T1Belief', ylbl='T2 Belief', data=d)

@@

```{r modelcompareplots, fig.width=11, fig.cap="Comparison of the coefficients and 95% CIs of the quasibinomial glm, fractional, and bootstrapped logistic regression models of T2 Belief and Action (for details, see the Statistical methods section)."}
